---
title: "Stop and Search data extraction"
author: Jolyon Miles-Wilson
output: html_document
---

This file documents the process for extracting Stop and Search data from the Police API

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(jsonify) # for reading json
library(leaflet) # for map
```

# Coordinates

The below code takes a json file specifying a geographic area and reads it into a data frame. Two data frames are created; one with column order longitude, latitude (for plotting on map), the other with column order latitude, longitude (for Police API).

```{r read_coords}
# Get coordinates of Birmingham area
file <- jsonify::from_json("../data/birm_geo.json") # read json
coords <- as.data.frame(file[["coordinates"]][[1]]) # make dataframe
colnames(coords) <- c("long","lat") # name cols
coords_for_map <- coords # map wants long, lat
coords <- subset(coords, select = c("lat","long")) # swap columns for API (wants lat, long)
```

The area specified by the above-defined coordinates captures the area below

```{r}
leaflet(as.matrix(coords_for_map)) %>% 
  addPolygons() %>% 
  addTiles()
```

# Data retrieval

The below function extracts stop and search data from the Police API based on a specified area and date range.

## Area retrieval
```{r area_data}
# Acquire stop and search data in an area for a time period

# area_coords: area of interest. Must be a data frame with column names "lat" and "long" 

# most_recent_month: numeric value specifying most recent month (e.g. 8 for August)

# most_recent_year (YYYY): numeric value specifying most recenet year (e.g. 2021)

# num_months_backwards: number of months backwards for which data are required. 
# Default = 12. For example, if the most recent year of data is required, 
# num_months_backwards would be 12

# If most_recent_month/year is not specified, function will query Police API for
# most recent update and use this as a start point

retrieve_area_data <- function(area_coords, most_recent_month = NULL, most_recent_year = NULL, num_months_backwards = 12){
  output <- data.frame() # initialise dataframe
  
  # make coordinate string for POST request
  coord_string <- c()
  for(i in 1:nrow(area_coords)){
    if(i == 1){
      coord_string <- paste(area_coords[i,"lat"],",",area_coords[i,"long"],sep="")
    }else{
    coord_string <- paste(coord_string,":",area_coords[i,"lat"],",",area_coords[i,"long"],sep="")
  }
  }
  # get most recent update if most_recent data not specified
  if(is.null(most_recent_month) || is.null(most_recent_year)){ 
    date <- httr::content(httr::GET("https://data.police.uk/api/crimes-street-dates"))[[1]][["date"]] # get most recent update
    most_recent_month <- as.numeric(substr(date,6,7))
    most_recent_year <- as.numeric(substr(date,1,4))
  }
  else{
    most_recent_month <- most_recent_month
    most_recent_year <- most_recent_year
  }
  
  # specify formatted date
  for(i in 1:num_months_backwards){
    if(i == 1){ # set values for first iteration
      month_num <- most_recent_month
      year <- most_recent_year
    }
    else{ # subsequent iterations
      month_num <- month_num - 1 # backwards a month each iteration
      if(month_num %% 12 == 0){ # if reach a new year, start months from 12 again
        month_num <- 12
        year <- year - 1 # backwards a year
      }
    }
    if(month_num < 10){ # paste 0 for months lower than 10
      month <- paste("0", month_num, sep = "")
    }
    else{
      month <- month_num
    }
    
    date <- paste(year, "-", month, sep = "") # combine dates into format for ukpolice
    body <- list("poly" = coord_string,
                 "date" = date)
    post_request <- httr::POST("https://data.police.uk/api/stops-street?", body = body) # POST request
    df <- httr::content(post_request) # get content from response
    df_2 <- data.frame(t(sapply(df,c))) # convert to df
    output <- rbind(output, df_2) # add to dataframe
    print(paste("Working... ", i, " of ", num_months_backwards, " retrieved", sep = ""))
    print(date)
  }
  # unlist remaining listed columns. Not ideal solution but will have to do for now
  output <- output %>%
    unnest_wider(., outcome_object, names_sep = ".") %>%
    unnest_wider(., location, names_sep = ".") %>%
    unnest_wider(., location.street, names_sep = ".")
  output <- as.matrix(output) # make matrix because write.csv runs into problems with list cols
  return(output) # testing... change back to output
}

data <- retrieve_area_data(coords, 8, 2021, 12)
write.csv(data, file = "../data/birmingham_stop_search_2020_09_2021_08.csv")
```

## Force retrieval
```{r force_list}
# Function for providing a list of forces for use with the API

force_list <- function(){
  list <- dplyr::bind_rows(httr::content(httr::GET("https://data.police.uk/api/forces")))
  list <- subset(list, select = c("name","id"))
  return(list)
}

#force_list()
```

```{r force_data}
# Acquire stop and search data by force for a time period

# force: force of interest (string). Available forces can be found using 
# force_list() function

# most_recent_month: numeric value specifying most recent month (e.g. 8 for August)

# most_recent_year (YYYY): numeric value specifying most recenet year (e.g. 2021)

# num_months_backwards: number of months backwards for which data are required. 
# Default = 12. For example, if the most recent year of data is required, 
# num_months_backwards would be 12

# If most_recent_month/year is not specified, function will query Police API for
# most recent update and use this as a start point

retrieve_force_data <- function(force, most_recent_month = NULL, most_recent_year = NULL, num_months_backwards = 12){
  output <- data.frame() # initialise dataframe
  if(is.null(most_recent_month) || is.null(most_recent_year)){ # if most_recent data not specified
    date <- httr::content(httr::GET("https://data.police.uk/api/crimes-street-dates"))[[1]][["date"]] # get most recent update
    most_recent_month <- as.numeric(substr(date,6,7))
    most_recent_year <- as.numeric(substr(date,1,4))
  }
  else{
    most_recent_month <- most_recent_month
    most_recent_year <- most_recent_year
  }
  for(i in 1:num_months_backwards){
    if(i == 1){ # set values for first iteration
      month_num <- most_recent_month
      year <- most_recent_year
    }
    else{ # subsequent iterations
      month_num <- month_num - 1 # backwards a month each iteration
      if(month_num %% 12 == 0){ # if reach a new year, start months from 12 again
        month_num <- 12
        year <- year - 1 # backwards a year
      }
    }
    if(month_num < 10){ # paste 0 for months lower than 10
      month <- paste("0", month_num, sep = "")
    }
    else{
      month <- month_num
    }
    
    date <- paste(year, "-", month, sep = "") # combine dates into format for ukpolice
    body <- list("force" = force,
                 "date" = date)
    post_request <- httr::POST("https://data.police.uk/api/stops-force?", body = body) # POST request
    df <- httr::content(post_request) # get content
    df_2 <- data.frame(t(sapply(df,c))) # convert to df
    output <- rbind(output, df_2) # run ukpolice and add to dataframe
    print(paste("Working... ", i, " of ", num_months_backwards, " retrieved", sep = ""))
    print(date)
  }
  return(output)
}

#stop_search_force_data <- retrieve_force_data("west-midlands", 8, 2021, 12) 
# last time force data compiled was March 2021
```







